Defocus Corrected Large Area Cryo-EM (DeCo-LACE) for Label-Free Detection of Molecules across Entire Cell Sections

### 2022-06-06 manuscript.tex
Incorporating more suggestions [ci skip]
This build is based on
https://github.com/jojoelfe/deco_lace_template_matching_manuscript/commit/a321def4b341f6e6f2c8d66936f53cf5e9398eae.

This commit was created by the following CI build and job:
https://github.com/jojoelfe/deco_lace_template_matching_manuscript/commit/a321def4b341f6e6f2c8d66936f53cf5e9398eae/checks
https://github.com/jojoelfe/deco_lace_template_matching_manuscript/actions/runs/2448512939

This manuscript
(\href{https://jojoelfe.github.io/deco_lace_template_matching_manuscript/v/26bf17f811e80363e8f38656b89bbae2c14147f8/}{permalink})
was automatically generated
from \href{https://github.com/jojoelfe/deco_lace_template_matching_manuscript/tree/26bf17f811e80363e8f38656b89bbae2c14147f8}{jojoelfe/deco\_lace\_template\_matching\_manuscript@26bf17f}
on May 27, 2022.

@@

\begin{itemize}
\item
  \textbf{Johannes Elferich}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.pdf}
  \href{https://orcid.org/0000-0002-9911-706X}{0000-0002-9911-706X}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.pdf}
  \href{https://github.com/jojoelfe}{jojoelfe}
  RNA Therapeutic Institute, UMass Chan Medical School; HHMI
\item
  \textbf{Giulia Schiroli}
  None
\item
  \textbf{David Scadden}
  None
\item
  \textbf{Nikolaus Grigorieff}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.pdf}
  \href{https://orcid.org/0000-0002-1506-909X}{0000-0002-1506-909X}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.pdf}
  \href{https://github.com/nikogrigorieff}{nikogrigorieff}
  RNA Therapeutic Institute, UMass Chan Medical School; HHMI
\end{itemize}

@@

Localization of biomolecules inside a cell is an important goal of biological
imaging. Fluorescence microscopy can localize biomolecules inside whole cells
and tissues, but its ability to count biomolecules and accuracy of the spatial
coordinates is limited by the wavelength of visible light. Cryo-electron
microscopy (cryo-EM) provides highly accurate position and orientation
information of biomolecules but is often confined to small fields of view inside
a cell, limiting biological context. In this study we use a new data-acquisition
scheme called ``Defocus-Corrected Large-Area cryo-EM'' (DeCo-LACE) to collect
high-resolution cryo-EM data over entire sections (100 -- 200 nm thick lamellae)
of neutrophil-like mouse cells, representing 1-2\% of the total cellular
volume. We use 2D template matching (2DTM) to determine localization and orientation
of the large ribosomal subunit in these sections. These data provide ``maps'' of translational activity across
sections of mammalian cells. This new high-throughput cryo-EM
data collection approach together with 2DTM will advance visual
proteomics and complement other single-cell ``omics'' techniques, such as
flow-cytometry and single-cell sequencing.

@@

A major goal in understanding cellular processes is the knowledge of the
amounts, location, interactions, and conformations of biomolecules
inside the cell. This knowledge can be obtained by approaches broadly
divided into label- and label-free techniques. In label-dependent
techniques a probe is physically attached to a molecule of interest that
is able to be detected by its strong signal, such as a
fluorescent molecule. In label-free techniques, the physical properties
of molecules themselves are used for detection. An example for this is
proteomics using mass-spectrometry {[}\protect\hyperlink{ref-tSXIKPl7}{1}{]}. The advantage
of label-free techniques is that they can provide information over
thousands of molecules, while label-dependent techniques offer highly specific
information for a few molecules. Especially spatial information can
often only be achieved using label-dependent techniques, such as
fluorescence microscopy {[}\protect\hyperlink{ref-VBmW7Aot}{2}{]}.

@@

Here we apply 2DTM of the ribosome large subunit (LSU) to cryo-FIB milled
neutrophil-like murine cells {[}\protect\hyperlink{ref-1B9Vt9eYu}{10}{]}. We chose these
cells because genetic defects in the ribosome machinery often leads to
hematopoietic disease {[}\protect\hyperlink{ref-gRoY21jY}{11}{]} and direct quantification of
ribosome location, number and conformational states in hematopoietic cells could
lead to new insight into hematopoietic disease {[}\protect\hyperlink{ref-KAJ7221k}{12}{]}.
To increase the amount of collected data and to provide unbiased sampling of the
whole lamella, we devised a new data-acquisition scheme, ``Defocus-Corrected
Large Area Cryo-Electron microscopy'' (DeCo-LACE). 2DTM allows us to test whether
aberrations caused by large beam-image shifts and highly condensed beams
deteriorate the high-resolution signal. We find that these aberrations do not
impede LSU detection by 2DTM. The resulting data provide a description of
ribosome distribution in an entire lamella, which represent 1-2\% of the cellular
volume. We find a highly heterogeneous density of ribosomes within the cell.
Analysis of the throughput in this method suggests that for the foreseeable
future computation will be the bottleneck for visual proteomics.

@@

To test whether we could detect individual ribosomes in mammalian cells we
prepared cryo-lamellae of mouse neutrophil-like cells. An overview image
of a representative lamella clearly shows cellular features consistent with a
neutrophil-like phenotype, mainly a segmented nucleus and a plethora of
membrane-organelles, corresponding to the granules and secretory vesicles of
neutrophils (Fig. {[}\ref{fig:initmatching}{]}A). We then proceeded to acquire
micrographs on this lamella with a defocus of 0.5-1.0 μm, 30 e\textsuperscript{-}/Å\(^2\)/s
exposure and 1.76 Å pixel size. We manually selected multiple locations in the
lamella and acquired micrographs using standard low-dose techniques where
focusing is performed on a sacrificial area. The resulting micrographs showed
smooth bilayered membranes and no signs of crystalline ice (Fig.
{[}\ref{fig:initmatching}{]}C,D), indicating successful vitrification throughout the
lamella. Successful vitrification is facilitated by the small size (\textasciitilde8 μm diameter) of these cells.

We used an atomic model of the 60S mouse ribosomal subunit (6SWA) for 2DTM {[}\protect\hyperlink{ref-sbkam02a}{13}{]}. In a
subset of images, the distribution of cross-correlation scores significantly
exceeded the distribution expected from images devoid of detectable targets. In the
resulting scaled maximum-intensity projections (MIPs), clear peaks with SNR
values up to 10 were apparent (Fig. {[}\ref{fig:initmatching2}{]}A). Using a
threshold criterion to select significant targets (see Methods), we found that in
images of cytosolic compartments there were 10-500 ribosomes within one micrograph
(Fig. {[}\ref{fig:initmatching}{]}B-E). Notably, we found no targets in areas
corresponding to the nucleus (Fig. {[}\ref{fig:initmatching}{]}B) or mitochondria
(Fig. \protect\hyperlink{fig:initmatching}{1}D). In the cytoplasm, we found a highly variable
number of targets, only \textasciitilde{} 50 in some exposures (Fig. {[}\ref{fig:initmatching}{]}E) and
up to 500 in others (Fig. {[}\ref{fig:initmatching}{]}C). This is a ten-fold
difference in local ribosome concentration, highlighting the importance of
imaging larger areas of a lamella to gain a complete picture of target
distributions. We therefore set out to collect cryo-EM data for 2DTM from
mammalian cell lamellae in a high-throughput unbiased fashion.

@@

In order to obtain high-resolution data from complete lamellae, we used a new
approach for data collection. This approach uses three key strategies: (1) every
electron that exposes a fresh area of the sample is collected on the camera (2)
image shift is used to precisely and quickly raster the surface of a lamella and
(3) focusing is done without using a sacrificial area (Fig. {[}\ref{fig:approach}{]}A).

To ensure that every electron exposing a fresh area of the sample is captured by
the detector, we adjusted the electron beam size to be entirely contained by the
detector area. During canonical low-dose imaging, the microscope is configured so
that the focal plane is identical to the eucentric plane of the specimen stage.
This leaves the C2 aperture out of focus, resulting in ripples at the edge of
the beam (Fig. {[}\ref{fig:approach}{]}D). While these ripples are low-resolution
features that likely do not interfere with 2DTM {[}cite{]}, we also tested data
collection under conditions where the C2 aperture is in focus (``fringe-free'',
Fig. {[}\ref{fig:approach}{]}E).

@@

We used this strategy to collect data on eight lamellae, four using the
eucentric focus condition, hereafter referred to as Lamella\textsubscript{EUC}, and four using
the fringe-free condition, hereafter referred to as Lamella\textsubscript{FFF}(Fig.
{[}\ref{fig:assembly}{]} A+D, Fig.
{[}\ref{fig:lamella_images}{]}A). We were able to collect data with a highly
consistent defocus of 800 nm (Fig. {[}\ref{fig:approach}{]}F), both in the eucentric
focus and fringe-free focus condition. To ensure that data were collected
consistently, we mapped defocus values as a function of the applied image shift
(Fig. {[}\ref{fig:lamella_spatial_info}{]}A). This demonstrated that the defocus was
consistent across a lamella, except for rare outliers and in images containing
contamination. We also plotted the measured objective astigmatism of each
lamella and found that it varies with the applied image shift, becoming more
astigmatic mostly due to image shift in the x direction (Fig.
{[}\ref{fig:lamella_spatial_info}{]}B). While approaches exist to correct for this
during the data collection {[}\protect\hyperlink{ref-APSL9LmU}{14}{]}, we opted to not use these approaches in our
initial experiments. We reasoned that because 2DTM depends on high-resolution
information, this would be an excellent test of how much these aberration affect
imaging.

@@

The montages show membrane vesicles and granules with highly variable sizes and
density. We found that a substantial number of granules, which are characterized by higher
density inside the the surrounding cytosol {[}\protect\hyperlink{ref-xE2EhAcH}{15}{]}, seemed to contain a membrane-enclosed inclusion with density similar to
the surrounding cytosol (Fig. {[}\ref{fig:lamella_images}{]}C) and could therefore be formed
by an autophagy-like pathway. These granules were
150-300 nm in diameter and the inclusions were 100-200 nm in diameter. Based on
these dimensions the granules are either azurophil or specific granules {[}\protect\hyperlink{ref-xE2EhAcH}{15}{]}. To our
knowledge, these inclusion have not been described in granulocytes and are
further described and discussed below.

\hypertarget{dtm-of-deco-lace-data-reveals-ribosome-distribution-in-cellular-cross-sections}{%
\subsubsection{2DTM of DeCo-LACE data reveals ribosome distribution in cellular cross-sections}\label{dtm-of-deco-lace-data-reveals-ribosome-distribution-in-cellular-cross-sections}}

In our initial attempts of using 2DTM on micrographs acquired with the DeCo-LACE
protocol, we did not observe any SNR peaks above threshold using the large
subunit of the mouse ribosome (Fig. {[}\ref{fig:crop_unblur}{]}A). We reasoned that
the edges of the beam might interfere with motion-correction of the movies as
they represent strong low-resolution features that do not move with the sample.
When we cropped the movie frames to exclude the beam edges, the estimated amount
of motion increased (Fig. {[}\ref{fig:crop_unblur}{]}B), consistent with successful
tracking of sample motion. Furthermore, in the motion-corrected average we could
identify significant SNR peaks (Fig. {[}\ref{fig:crop_unblur}{]}B), confirming to
high sensitivity of 2DTM to the presents of high-resolution signal preserved in
the images by the motion correction. To streamline data processing, we
implemented a function in unblur to consider only a defined central area of a
movie for estimation of sample motion, while still averaging the complete movie
frames (Fig. {[}\ref{fig:crop_unblur}{]}C). Using this approach, we motion-corrected
all tiles in the eight lamellae and found consistently total motion below 1 Å
per frame (Fig. {[}\ref{fig:lamella_motion_thickness}{]} A). In some lamellae we
found increased motion in the lamella center, which indicates areas of variable
mechanical stability within FIB-milled lamellae. In some micrographs we also
observed that the beam edges gave rise to artifacts in the MIP and numerous
false-positive detections at the edge of the illuminated area (Fig. {[}\ref{fig:crop_unblur}{]}D). A
similar phenomenon was observed on isolated ``hot'' pixels in unilluminated areas.
To overcome this issue we implemented a function in unblur to replace
dark areas in the micrograph with Gaussian noise (see Methods), with
mean and standard deviation matching the illuminated portion of the micrograph
(Fig. {[}\ref{fig:crop_unblur}{]}D+E). Together, these pre-processing steps enabled us to perform 2DTM on
all tiles of the eight lamellae.

We used the refined tile positions to calculate the positions of the detected
LSUs in the lamellae (Fig. {[}\ref{fig:matching_euc}{]}A, Fig. {[}\ref{fig:matching_fff}{]}A). Overlaying these positions of the lamellae montages reveal ribosome
distribution throughout the FIB-milled slices of individual cells. Organelles
like the nucleus and mitochondria only showed sporadic targets detected with low
SNRs, consistent with the estimated false-positive rate of one per tile. For
each detected target we also calculated the Z positions from the individual
estimated defocus and defocus offset for each tile. When viewed from the side,
the ribosome positions therefore show the slight tilts of the lamellae relative
to the microscope frame of reference (Fig. {[}\ref{fig:matching_euc}{]}B, Fig.
{[}\ref{fig:matching_fff}{]}B). Furthermore, the side views indicated that lamellae
were thinner at the leading edge. Indeed, when plotting the transmitted beam
intensities in individual tiles as a function of beam image-shift, we observed
substantially higher intensities at the leading edge (Fig. {[}\ref{fig:lamella_motion_thickness}{]}B), which in energy-filtered
TEM indicates a thinner sample {[}\protect\hyperlink{ref-17IP4Lhz1}{16}{]}. Even though we prepared the lamellae
with the ``overtilt'' approach {[}\protect\hyperlink{ref-hr3dyUeI}{17}{]}, this means that ribosome densities across
the lamellae can be skewed by a change in thickness, and better sample
preparation methods are needed to generate more even samples.

Close inspection of the ribosome positions in the lamellae revealed several
interesting features. Ribosomes could be seen associating with membranes, in
patterns reminiscent of the rough endoplasmic reticulum (Fig.
{[}\ref{fig:matching_euc}{]}C, Fig. {[}\ref{fig:matching_fff}{]}C) or the outer nuclear
membrane (Fig. {[}\ref{fig:matching_euc}{]}D). We also observed ribosomes forming
ring-like structures (Fig. {[}\ref{fig:matching_euc}{]}E), potentially indicating
circularized mRNAs {[}\protect\hyperlink{ref-cie13Q8F}{18}{]}. While ribosomes were for the most part excluded
from the numerous granules observed in the cytoplasm, in some cases we observed
clusters of ribosomes in the inclusions of double-membraned granules
described earlier (Fig. {[}\ref{fig:matching_euc}{]}F, Fig. {[}\ref{fig:matching_fff}{]}D,E). It is, in principle, possible that these targets are situated above or
below the imaged granules, since the granule positions in z cannot be determined
using 2D projections. However, in the case of Fig. {[}\ref{fig:matching_fff}{]}E, the
detected ribosomes span the whole lamella in the z direction (Fig. {[}\ref{fig:matching_fff}{]}F), while positions
above or below a granule would result in ribosomes situated exlusively at the
top or bottom of the lamella. This conclusive evidence of ribosomes in
inclusions is consistent with the earlier hypothesis that the inclusions are of
cytoplasmic origin.

@@

Within the eight lamellae we found different number of detected targets (Fig.
{[}\ref{fig:matching_stat}{]}A). Lamella\textsubscript{EUC} 1 had the most detected targets, but also has
the largest surface area and contained cytoplasm from two cells. Lamella\textsubscript{FFF} 4
had the fewest detected targets, but this particular lamella was dominated by a
circular section of the nucleus, with only small pockets of cytoplasm (Fig.
{[}\ref{fig:lamella_images}{]}). In an attempt to normalize for these differences in
area containing cytoplasm, we compared the number of detected targets per tile in
tiles that contained more than one target, which should exclude tiles with
non-cytosolic content (Fig. {[}\ref{fig:matching_stat}{]}B). While this measure had
less variability, there were still differences. Lamella\textsubscript{EUC} 4 had not only the
fewest targets, but also the lowest density, which could be due to this lamella being the
thinnest, or due to it sectioning the cell in an area with a lower concentration of
ribosomes. Lamella\textsubscript{FFF} 3 had a substantially higher number of ribosomes per
tile. Since all of these lamellae were made from a cell-line under identical
conditions, this underscores the necessity to collect data from large numbers of
lamellae to overcome the inherent variability. When comparing the distribution of
scores between lamellae, we found them to be fairly comparable with median
SNRs ranging from 8.7 to 9.7 (Fig. {[}\ref{fig:matching_stat}{]}C). Lamella\textsubscript{EUC} 1 had slightly lower scores compared
to
the rest, potentially due to its large size and connected mechanical instability
during imaging. Overall, we did not observe differences in the number or SNR of
detected targets between eucentric or fringe-free illumination conditions that
were bigger than the observed inter-lamella variability.

@@

A list of tile pairs \(i,j\) that overlap were assembled by selecting images where
\(|\mathbf{c}_i-\mathbf{c}_j| < D_{Beam}\). In order to calculate the precise offset between tiles \(i\) and \(j\), \(\mathbf{r}_{i,j}\), we calculated the cross-correlation between the two tiles, masked to the overlapping illuminated area using the scikit-image
package {[}\protect\hyperlink{ref-stvWEJeu}{33}{]} was used to calculate refined offsets . The coordinates \(\mathbf{c}_{i}\) were then refined by a least-square minimization against \(\mathbf{r}_{i,j}\):

@@

>>>>

This manuscript
(\href{https://jojoelfe.github.io/deco_lace_template_matching_manuscript/v/a321def4b341f6e6f2c8d66936f53cf5e9398eae/}{permalink})
was automatically generated
from \href{https://github.com/jojoelfe/deco_lace_template_matching_manuscript/tree/a321def4b341f6e6f2c8d66936f53cf5e9398eae}{jojoelfe/deco\_lace\_template\_matching\_manuscript@a321def}
on June 6, 2022.

@@

\begin{itemize}
\item
  \textbf{Johannes Elferich}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.pdf}
  \href{https://orcid.org/0000-0002-9911-706X}{0000-0002-9911-706X}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.pdf}
  \href{https://github.com/jojoelfe}{jojoelfe}
  RNA Therapeutics Institute, University of Massachusetts Chan Medical School, Worcester, MA, USA; Howard Hughes Medical Institute, Ashburn, VA, USA
\item
  \textbf{Giulia Schiroli}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.pdf}
  \href{https://orcid.org/0000-0002-9812-8728}{0000-0002-9812-8728}
  Center for Regenerative Medicine, Massachusetts General Hospital, Boston, MA 02114, USA
\item
  \textbf{David Scadden}
  Center for Regenerative Medicine, Massachusetts General Hospital, Boston, MA 02114, USA; Harvard Stem Cell Institute, Cambridge, MA 02138, USA; Department of Stem Cell and Regenerative Biology, Harvard University, Cambridge, MA 02138, USA
\item
  \textbf{Nikolaus Grigorieff}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.pdf}
  \href{https://orcid.org/0000-0002-1506-909X}{0000-0002-1506-909X}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.pdf}
  \href{https://github.com/nikogrigorieff}{nikogrigorieff}
  RNA Therapeutics Institute, University of Massachusetts Chan Medical School, Worcester, MA, USA; Howard Hughes Medical Institute, Ashburn, VA, USA
\end{itemize}

@@

A major goal of biological imaging is localization of biomolecules inside a cell. Fluorescence microscopy can localize biomolecules inside whole cells
and tissues, but its ability to count biomolecules and accuracy of the spatial
coordinates is limited by the wavelength of visible light. Cryo-electron
microscopy (cryo-EM) provides highly accurate position and orientation
information of biomolecules but is often confined to small fields of view inside
a cell, limiting biological context. In this study we use a new data-acquisition
scheme called ``Defocus-Corrected Large-Area cryo-EM'' (DeCo-LACE) to collect
high-resolution images of entire sections (100 -- 200 nm thick lamellae)
of neutrophil-like mouse cells, representing 1-2\% of the total cellular
volume. We use 2D template matching (2DTM) to determine localization and orientation of the large ribosomal subunit in these sections. These data provide ``maps'' of ribosomes across entire sections of mammalian cells. This high-throughput cryo-EM data collection approach together with 2DTM will advance visual proteomics and provide biological insight that cannot be obtained by other methods.

@@

A major goal in understanding cellular processes is the knowledge of the
amounts, location, interactions, and conformations of biomolecules
inside the cell. This knowledge can be obtained by approaches broadly
divided into label- and label-free techniques. In label-dependent
techniques a probe is physically attached to a molecule of interest that
is able to be detected by its strong signal, such as a
fluorescent molecule. In label-free techniques, the physical properties
of molecules themselves are used for detection. An example for this is
proteomics using mass-spectrometry {[}\protect\hyperlink{ref-tSXIKPl7}{1}{]}. The advantage
of label-free techniques is that they can provide information over
thousands of molecules, while label-dependent techniques offer highly specific
information for a few molecules. in particular, spatial information is primarily achieved using label-dependent techniques, such as
fluorescence microscopy {[}\protect\hyperlink{ref-VBmW7Aot}{2}{]}.

@@

Here we apply 2DTM of the ribosome large subunit (LSU) to conditionally immortalized \emph{mus musculus} (mouse) cell line that gives rise to functional mature neutrophils {[}\protect\hyperlink{ref-1B9Vt9eYu}{10}{]} after thinning using cryo-FIB milling. We chose these
cells because genetic defects in the ribosome machinery often leads to
hematopoietic disease {[}\protect\hyperlink{ref-gRoY21jY}{11}{]} and direct quantification of
ribosome location, number and conformational states in hematopoietic cells could
lead to new insight into hematopoietic disease {[}\protect\hyperlink{ref-KAJ7221k}{12}{]}.
To increase the amount of collected data and to provide unbiased sampling of the
whole lamella, we devised a new data-acquisition scheme, ``Defocus-Corrected
Large Area Cryo-Electron microscopy'' (DeCo-LACE). 2DTM allows us to test whether
aberrations caused by large beam-image shifts and highly condensed beams
deteriorate the high-resolution signal. We find that these aberrations do not
impede LSU detection by 2DTM. The resulting data provide a description of
ribosome distribution in an entire lamella, which represent 1-2\% of the cellular
volume. We find a highly heterogeneous density of ribosomes within the cell.
Analysis of the throughput in this method suggests that for the foreseeable
future computation will be the bottleneck for visual proteomics.

@@

FIB-milled \emph{Saccharomyces cerevisiae} (yeast) cells are sufficiently well preserved to permit localization of 60S ribosomal subunits with 2DTM {[}\protect\hyperlink{ref-T1Qr0xyP}{13}{]}. Due to the larger size of mammalian cells compared to yeast cells, it was unclear whether plunge freezing would be adequate to produce vitreous ice across the whole volume of the cell. To test this we prepared cryo-lamellae of mouse neutrophil cells. A low magnification image
of a representative lamella clearly shows cellular features consistent with a
neutrophile-like phenotype, mainly a segmented nucleus and a plethora of
membrane-organelles, corresponding to the granules and secretory vesicles of
neutrophils (Fig. {[}\ref{fig:initmatching}{]}A). We then proceeded to acquire
micrographs on this lamella with a defocus of 0.5-1.0 μm, 30 e\textsuperscript{-}/Å\(^2\)/s
exposure and 1.76 Å pixel size. We manually selected multiple locations in the
lamella and acquired micrographs using standard low-dose techniques where
focusing is performed on a sacrificial area. The resulting micrographs showed
smooth bilayered membranes and no signs of crystalline ice (Fig.
{[}\ref{fig:initmatching}{]}C,D), indicating successful vitrification throughout the
lamella.

We used an atomic model of the 60S mouse ribosomal subunit (6SWA) for 2DTM {[}\protect\hyperlink{ref-sbkam02a}{14}{]}. In a
subset of images, the distribution of cross-correlation scores significantly
exceeded the distribution expected from images devoid of detectable targets. In the
resulting scaled maximum-intensity projections (MIPs), clear peaks with SNR
values up to 10 were apparent (Fig. {[}\ref{fig:initmatching2}{]}A). Using a
threshold criterion to select significant targets (see Methods), we found that in
images of cytosolic compartments there were 10-500 ribosomes within one micrograph
(Fig. {[}\ref{fig:initmatching}{]}B-E). Notably, we found no targets in areas
corresponding to the nucleus (Fig. {[}\ref{fig:initmatching}{]}B) or mitochondria
(Fig. \protect\hyperlink{fig:initmatching}{1}D). In the cytoplasm, we found a highly variable
number of targets, only \textasciitilde{} 50 in some exposures (Fig. {[}\ref{fig:initmatching}{]}E) and
up to 500 in others (Fig. {[}\ref{fig:initmatching}{]}C). However, it is unclear whether this ten-fold difference in local ribosome concentration is due to technical variation, such as sample thickness, or biological variation. To differentiate between the two we reasoned it was important to not manually choose imaging regions and to collect larger amounts of data. We therefore set out to collect cryo-EM data for 2DTM from mammalian cell lamellae in a high-throughput unbiased fashion.

@@

In order to obtain high-resolution data from complete lamellae, we developed a new
approach for data collection. This approach uses three key strategies: (1) every
electron that exposes a fresh area of the sample is collected on the camera (2)
image shift is used to precisely and quickly raster the surface of a lamella and
(3) focusing is done without using a sacrificial area (Fig. {[}\ref{fig:approach}{]}A).

To ensure that every electron exposing a fresh area of the sample is captured by
the detector, we adjusted the electron beam size to be entirely contained by the
detector area. During canonical low-dose imaging, the microscope is configured so
that the focal plane is identical to the eucentric plane of the specimen stage.
This leaves the C2 aperture out of focus, resulting in ripples at the edge of
the beam (Fig. {[}\ref{fig:approach}{]}D). While these ripples are low-resolution
features that likely do not interfere with 2DTM {[}\protect\hyperlink{ref-Ynb3IP6I}{7}{]}, we also tested data
collection under conditions where the C2 aperture is in focus (``fringe-free'',
Fig. {[}\ref{fig:approach}{]}E) {[}\protect\hyperlink{ref-wsFP5c54}{15}{]}.

@@

We used this strategy to collect data on eight lamellae, four using the
eucentric focus condition, hereafter referred to as Lamella\textsubscript{EUC}, and four using
the fringe-free condition, hereafter referred to as Lamella\textsubscript{FFF}(Fig.
{[}\ref{fig:assembly}{]} A+D, Fig.
{[}\ref{fig:lamella_images}{]}A). We were able to collect data with a highly
consistent defocus of 800 nm (Fig. {[}\ref{fig:approach}{]}F), both in the eucentric
focus and fringe-free focus condition. To ensure that data were collected
consistently, we mapped defocus values as a function of the applied image shift
(Fig. {[}\ref{fig:lamella_spatial_info}{]}A). This demonstrated that the defocus was
consistent across a lamella, except for rare outliers and in images containing
contamination. We also plotted the measured objective astigmatism of each
lamella and found that it varies with the applied image shift, becoming more
astigmatic mostly due to image shift in the x direction (Fig.
{[}\ref{fig:lamella_spatial_info}{]}B). While approaches exist to correct for this
during the data collection {[}\protect\hyperlink{ref-APSL9LmU}{16}{]}, we opted to not use these approaches in our
initial experiments. We reasoned that because 2DTM depends on high-resolution
information, this would be an excellent test of how much these aberration affect
imaging.

@@

The montages show membrane vesicles and granules with highly variable sizes and
density. We found that a substantial number of granules, which are characterized by higher
density inside the the surrounding cytosol {[}\protect\hyperlink{ref-xE2EhAcH}{17}{]}, seemed to contain a membrane-enclosed inclusion with density similar to
the surrounding cytosol (Fig. {[}\ref{fig:lamella_images}{]}C) and could therefore be formed
by inward budding of the granule membrane. These granules were
150-300 nm in diameter and the inclusions were 100-200 nm in diameter. Based on
these dimensions the granules are either azurophil or specific granules {[}\protect\hyperlink{ref-xE2EhAcH}{17}{]}. To our
knowledge, these inclusion have not been described in granulocytes and are
further described and discussed below.

\hypertarget{dtm-of-deco-lace-data-reveals-large-ribosomal-subunit-distribution-in-cellular-cross-sections}{%
\subsubsection{2DTM of DeCo-LACE data reveals large ribosomal subunit distribution in cellular cross-sections}\label{dtm-of-deco-lace-data-reveals-large-ribosomal-subunit-distribution-in-cellular-cross-sections}}

In our initial attempts of using 2DTM on micrographs acquired with the DeCo-LACE
protocol, we did not observe any SNR peaks above threshold using the large
subunit of the mouse ribosome (Fig. {[}\ref{fig:crop_unblur}{]}A). We reasoned that
the edges of the beam might interfere with motion-correction of the movies as
they represent strong low-resolution features that do not move with the sample.
When we cropped the movie frames to exclude the beam edges, the estimated amount
of motion increased (Fig. {[}\ref{fig:crop_unblur}{]}B), consistent with successful
tracking of sample motion. Furthermore, in the motion-corrected average we could
identify significant SNR peaks (Fig. {[}\ref{fig:crop_unblur}{]}B), confirming the
high sensitivity of 2DTM to the presence of high-resolution signal preserved in
the images by the motion correction. To streamline data processing, we
implemented a function in unblur to consider only a defined central area of a
movie for estimation of sample motion, while still averaging the complete movie
frames (Fig. {[}\ref{fig:crop_unblur}{]}C). Using this approach, we motion-corrected
all tiles in the eight lamellae and found consistently total motion below 1 Å
per frame (Fig. {[}\ref{fig:lamella_motion_thickness}{]} A). In some lamellae we
found increased motion in the lamella center, which indicates areas of variable
mechanical stability within FIB-milled lamellae. In some micrographs we also
observed that the beam edges gave rise to artifacts in the MIP and numerous
false-positive detections at the edge of the illuminated area (Fig. {[}\ref{fig:crop_unblur}{]}D). A
similar phenomenon was observed on isolated ``hot'' pixels in unilluminated areas.
To overcome this issue we implemented a function in unblur to replace
dark areas in the micrograph with Gaussian noise (see Methods), with
mean and standard deviation matching the illuminated portion of the micrograph
(Fig. {[}\ref{fig:crop_unblur}{]}D+E). Together, these pre-processing steps enabled us to perform 2DTM on
all tiles of the eight lamellae.

We used the tile positions to calculate the positions of the detected
LSUs in the lamellae (Fig. {[}\ref{fig:matching_euc}{]}A, Fig. {[}\ref{fig:matching_fff}{]}A). Overlaying these positions of the lamellae montages reveals LSU
distribution throughout the FIB-milled slices of individual cells. Consistent with prior observations imaging selected views in yeast {[}\protect\hyperlink{ref-T1Qr0xyP}{13}{]}, organelles
like the nucleus and mitochondria only showed sporadic targets detected with low
SNRs, consistent with the estimated false-positive rate of one per tile. For
each detected target we also calculated the Z positions from the individual
estimated defocus and defocus offset for each tile. When viewed from the side,
the ribosome positions therefore show the slight tilts of the lamellae relative
to the microscope frame of reference (Fig. {[}\ref{fig:matching_euc}{]}B, Fig.
{[}\ref{fig:matching_fff}{]}B). Furthermore, the side views indicated that lamellae
were thinner at the leading edge. Indeed, when plotting the transmitted beam
intensities in individual tiles as a function of beam image-shift, we observed
substantially higher intensities at the leading edge (Fig. {[}\ref{fig:lamella_motion_thickness}{]}B), which in energy-filtered
TEM indicates a thinner sample {[}\protect\hyperlink{ref-17IP4Lhz1}{18}{]}. Even though we prepared the lamellae
with the ``overtilt'' approach {[}\protect\hyperlink{ref-hr3dyUeI}{19}{]}, this means that LSU densities across
the lamellae can be skewed by a change in thickness, and better sample
preparation methods are needed to generate more even samples.

Close inspection of the LSU positions in the lamellae revealed several
interesting features. LSUs could be seen associating with membranes, in
patterns reminiscent of the rough endoplasmic reticulum (Fig.
{[}\ref{fig:matching_euc}{]}C, Fig. {[}\ref{fig:matching_fff}{]}C) or the outer nuclear
membrane (Fig. {[}\ref{fig:matching_euc}{]}D). We also observed LSUs forming
ring-like structures (Fig. {[}\ref{fig:matching_euc}{]}E), potentially indicating
circularized mRNAs {[}\protect\hyperlink{ref-cie13Q8F}{20}{]}. While ribosomes were for the most part excluded
from the numerous granules observed in the cytoplasm, in some cases we observed
clusters of LSUs in the inclusions of double-membraned granules
described earlier (Fig. {[}\ref{fig:matching_euc}{]}F, Fig. {[}\ref{fig:matching_fff}{]}D,E). It is, in principle, possible that these targets are situated above or
below the imaged granules, since the granule positions in z cannot be determined
using 2D projections. However, in the case of Fig. {[}\ref{fig:matching_fff}{]}E, the
detected LSUs span the whole lamella in the z direction (Fig. {[}\ref{fig:matching_fff}{]}F), while positions
above or below a granule would result in LSUs situated exlusively at the
top or bottom of the lamella. This is consistent with the earlier hypothesis that the inclusions are of
cytoplasmic origin.

@@

Within the eight lamellae we found different numbers of detected targets, ranging from 1089 to 6433 per lamella (Fig.
{[}\ref{fig:matching_stat}{]}A). Lamella\textsubscript{EUC} 1 had the most detected targets, but also has
the largest surface area and contained cytoplasm from two cells. Lamella\textsubscript{FFF} 4
had the fewest detected targets, but this particular lamella was dominated by a
circular section of the nucleus, with only small pockets of cytoplasm (Fig.
{[}\ref{fig:lamella_images}{]}). In an attempt to normalize for these differences in
area containing cytoplasm, we compared the number of detected targets per tile in
tiles that contained more than one target, which should exclude tiles with
non-cytosolic content (Fig. {[}\ref{fig:matching_stat}{]}B). While this measure had
less variability, there were still differences. Lamella\textsubscript{EUC} 4 had not only the
fewest targets, but also the lowest density, which could be due to this lamella being the
thinnest, or due to it sectioning the cell in an area with a lower concentration of
ribosomes. Lamella\textsubscript{FFF} 3 had a substantially higher number of ribosomes per
tile. Since all of these lamellae were made from a cell-line under identical
conditions, this underscores the necessity to collect data from large numbers of
lamellae to overcome the inherent variability. When comparing the distribution of
scores between lamellae, we found them to be fairly comparable with median
SNRs ranging from 8.7 to 9.7 (Fig. {[}\ref{fig:matching_stat}{]}C). Lamella\textsubscript{EUC} 1 had slightly lower scores compared
to
the rest, potentially due to its large size and connected mechanical instability
during imaging. Overall, we did not observe differences in the number or SNR of
detected targets between eucentric or fringe-free illumination conditions that
were bigger than the observed inter-lamella variability.

@@

A list of tile pairs \(i,j\) that overlap were assembled by selecting images where
\(|\mathbf{c}_i-\mathbf{c}_j| < D_{Beam}\). In order to calculate the precise offset between tiles \(i\) and \(j\), \(\mathbf{r}_{i,j}\), we calculated the cross-correlation between the two tiles, masked to the overlapping illuminated area using the scikit-image
package {[}\protect\hyperlink{ref-stvWEJeu}{34}{]} was used to calculate refined offsets . The coordinates \(\mathbf{c}_{i}\) were then refined by a least-square minimization against \(\mathbf{r}_{i,j}\):

@@

\begin{fignos:tagged-figure}[Movie 1]

\begin{figure}
\hypertarget{fig:movie1}{%
\centering
\includegraphics{images/1x1-00000000.png}
\caption{Movie of detected LSU targets in Lamella\(_\textrm{EUC}\) 1, corresponding to Figure 5}\label{fig:movie1}
}
\end{figure}
